---
title: "Marine Debris by Home States"
author: "Katharine Egan"
output: html_document
---

Libraries and functions:
```{r}
library(tidyverse)
library(readxl)

# read in functions needed to clean data 
source("scripts/functions/clean_summary_data.R")
source("scripts/functions/clean_ppm_data.R")
source("scripts/functions/read_in_rawdata.R")
```

Reading in data and functions:
```{r}
# read in summary of marine debris data
read_in_rawdata(rawdata_path = "raw_data/summary/", 
                list_name = "summary_list") 

# ppm data
read_in_rawdata(rawdata_path = "raw_data/ppm/", 
                list_name = "ppm_list") 
```

Clean up marine debris categories:

Note: 
-The Ocean Conservancy datacard changes from 2012 to 2013 so the trash categories are different hence writing out two separate data frames.
```{r}
# clean the list and make into a giant cleaned up data frame
# assigns two data frames to global environment
clean_summary_data(summary_list)
```

Cleanup marine debris metrics:
```{r}
# clean up PPM data (people, pounds, miles, # of bags)
# also split up into two separate data frames
clean_ppm_data(ppm_list)
```

Aggregate marine debris data by year and location (sum the total amount of debris), join the ppm data to the aggregated data frame, calculate debris density (pounds per mile), and calculate debris density per person (pounds per mile per person)
```{r}
# 2008-2012
md_ppm_0812 <- md_0812 %>% group_by(year, location) %>% 
  summarise(total_sum = sum(total)) %>% 
  full_join(ppm_0812) %>% 
  mutate(debris_density = total_sum/miles,
         debris_density_person = debris_density/people)

# 2013-2015
md_ppm_1315 <- md_1315 %>% group_by(year, location) %>% 
  summarise(total_sum = sum(total)) %>% 
  full_join(ppm_1315) %>% 
  mutate(debris_density = total_sum/miles,
         debris_density_person = debris_density/people)
```

