---
title: "Marine Debris by Home States or Territories"
subtitle: "Ocean Conservancy International Coastal Cleanup (2008-2015)"
author: "Katharine Egan"
output: html_document
---

Libraries and functions:
```{r}
library(tidyverse)
library(readxl)
library(cowplot)
library(scales)
library(RColorBrewer)

# read in functions needed to clean data 
source("scripts/functions/clean_summary_data.R")
source("scripts/functions/clean_ppm_data.R")
source("scripts/functions/read_in_rawdata.R")
```

Reading in data and functions:
```{r}
# read in summary of marine debris data
read_in_rawdata(rawdata_path = "raw_data/summary/", 
                list_name = "summary_list") 

# ppm data
read_in_rawdata(rawdata_path = "raw_data/ppm/", 
                list_name = "ppm_list") 
```

Clean up marine debris categories:

Note: 
-The Ocean Conservancy datacard changes from 2012 to 2013 so the trash categories are different hence writing out two separate data frames.
```{r}
# clean the list and make into a giant cleaned up data frame
# assigns two data frames to global environment
clean_summary_data(summary_list)
```

Cleanup marine debris metrics:
```{r}
# clean up PPM data (people, pounds, miles, # of bags)
# also split up into two separate data frames
clean_ppm_data(ppm_list)
```

Exploratory data analysis:
-Debris density per person by location and year
```{r}
# join all debris data with the PPM data 
# make a kilograms and kilometer column (from pounds and miles)
# calculate debris density (kg/km) 
# calculate debris density per person (kg/km/person)
md_ppm <- all_md %>% group_by(year, location) %>% 
  summarise(total_sum = sum(total)) %>% 
  full_join(all_ppm) %>% 
  mutate(kg = pounds * 0.453592,
         km = miles * 1.60934,
         debris_density_kgkm = kg/km,
         debris_density_lbsmi = pounds/miles,
         pieces_mile = total_sum/miles)

# graph of debris density (pounds/mile) by location (boxplot)
dd_lbsmi_loc <- ggplot(data = md_ppm,
       aes(y=debris_density_lbsmi, x=location)) +
  geom_boxplot(fill = "#80CDC1") +
  geom_jitter(size = 2) +
  scale_x_discrete(limits = c("PA", "RI", "NC", "USVI")) +
  labs(y = "Debris Density (lbs/mile)",
       x = "Home State/Territory") +
  theme_classic() +
  theme(text = element_text(size=12)) 

# graph of pieces per mile by location (boxplot)
dd_piecemi_loc <- ggplot(data = md_ppm,
       aes(y=pieces_mile, x=location)) +
  geom_boxplot(fill = "#8C510A") +
  geom_jitter(size = 2) +
  scale_x_discrete(limits = c("PA", "RI", "NC", "USVI")) +
  labs(y = "Debris Per Mile",
       x = "Home State/Territory") +
  theme_classic() +
  theme(text = element_text(size=12)) 

# make a grid plot of debris density graphs 
density_plot <- plot_grid(dd_lbsmi_loc, dd_piecemi_loc, align = "h", nrow = 1)

# summarized data of both types of debris density by location 
debris_location <- md_ppm %>% group_by(location) %>% 
  summarise(mean_lbsmi = mean(debris_density_lbsmi),
            sd_lbsmi = sd(debris_density_lbsmi),
            count_lbsmi = length(debris_density_lbsmi),
            se_lbsmi = sd(debris_density_lbsmi)/
              sqrt(length(debris_density_lbsmi)),
            mean_piecesmi = mean(pieces_mile),
            sd_piecesmi = sd(pieces_mile),
            count_piecesmi = length(pieces_mile),
            se_piecesmi = sd(pieces_mile)/
              sqrt(length(pieces_mile)))
```

Graphs of people, pounds, pieces of debris, and miles by home state/territory:
```{r}
# people
people <- ggplot(data = md_ppm,
       aes(y=people, x=location)) +
  geom_boxplot(fill = "#D7191C") +
  labs(y = "People") +
  scale_x_discrete(limits = c("PA", "RI","NC", "USVI")) +
  scale_y_continuous(labels = comma) +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_text(vjust=-0.8),
        text = element_text(size=12))

# pounds
pounds <- ggplot(data = md_ppm,
       aes(y=pounds, x=location)) +
  geom_boxplot(fill = "#FDAE61") +
  labs(y = "Pounds") +
  scale_x_discrete(limits = c("PA", "RI","NC", "USVI")) +
  scale_y_continuous(labels = comma) +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_text(vjust=-0.8),
        text = element_text(size=12))

# miles
miles <- ggplot(data = md_ppm,
       aes(y=miles, x=location)) +
  geom_boxplot(fill = "#ABDDA4") +
  labs(y = "Miles") +
  scale_x_discrete(limits = c("PA", "RI","NC", "USVI")) +
  scale_y_continuous(labels = comma) +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_text(vjust=-0.8),
        text = element_text(size=12))

# pieces of debris
debris <- ggplot(data = md_ppm,
       aes(y=total_sum, x=location)) +
  geom_boxplot(fill = "#2B83BA") +
  labs(y = "Total Debris Pieces") +
  scale_x_discrete(limits = c("PA", "RI","NC", "USVI")) +
  scale_y_continuous(labels = comma) +
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_text(vjust=-0.8),
        text = element_text(size=12))

# make one graph of all parts 
all_md_locations <- plot_grid(people, pounds, miles, debris, ncol = 2,
                          align = "vh")
all_md_locations
```

Top 5 items found in each state or territory across all years: 
```{r}
# get top 5 items across 2008-2012
top5_0812 <- md_0812 %>%
  group_by(location, category) %>%
  summarize(total = sum(total)) %>% 
  arrange(desc(total)) %>%
  slice(1:5) %>% 
  mutate(years = "2008-2012")

# get top 5 items across 2013-2015
top5_1315 <- md_1315 %>%
  group_by(location, category) %>%
  summarize(total = sum(total)) %>% 
  arrange(desc(total)) %>%
  slice(1:5) %>% 
  mutate(years = "2013-2015")
```
